Towards atomistic resolution structure of phosphatidylcholine headgroup and glycerol backbone at different ambient conditions\footnote{Publication about results presented in the NMRlipids project}

### 2015-08-09 HGmodelMANUSCRIPT/REPLYtoREVIEWER.txt

REPLY TO 1: We were trying to resolve the atomistic resolution structure of glycerol backbone 
and choline by using the available atomistic resolution models. In all of the used models the 
most recent version published by the developers was intended to be used in NPT ensemble with 
zero tension, indicating that the general opinion in the field consider the zero tension ensemble 
being more realistic. Most notably, the two popular lipid force fields that required the use of either 
a fixed membrane area or a nonzero surface tension, Charmm27 and Lipid11, were recently updated 
to Charmm36 and Lipid14, which removed this requirement. The older versions of the force fields
ment to be ran in non-zero tension are not discussed in the manuscript. Thus, the more detailed discussion
on this issue is beyond the scope of the work. Also, the improvements of the models and testing 
of influence of changes in parameters, including the ensemble, is beyond the scope of the work. 
Further, it is not very likely that the change in the ensemble would significantly change the 
glycerol backbone or choline structure. We have clarified that we only reproduce the simulations
with original parametrizations by the sentence: "All simulation parameters were set as close to the original 
parametrization works as possible." in the page 10.

### 2015-08-12 HGmodelMANUSCRIPT/HGmodel_ACSsupplementary.tex
Add glycerol orientation analysis and discussion to the manuscript, related to the reviewer's point 6

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Place any additional packages needed here.  Only include packages
%% which are essential, to avoid problems later. Do NOT use any
%% packages which require e-TeX (for example etoolbox): the e-TeX
%% extensions are not currently available on the ACS conversion
%% servers.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage[version=3]{mhchem} % Formula subscripts using \ce{}
\usepackage[obeyFinal]{easy-todo}
\usepackage{rotating}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% If issues arise when submitting your manuscript, you may want to
%% un-comment the next line.  This provides information on the
%% version of every file you have used.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%\listfiles

@@

\section{Dihedral angle distributions as a function of cholesterol in CHARMM36}
\begin{figure*}[]
  \centering
  \includegraphics[width=17.2cm]{../Fig/dihsCHOLcharmm.pdf}
  \caption{\label{dihsCHOLcharmm}
    The effect of cholesterol content on the POPC glycerol backbone and choline dihedral angles in CHARMM36 model (T=303~K).}
\end{figure*}

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Place any additional packages needed here.  Only include packages
%% which are essential, to avoid problems later. Do NOT use any
%% packages which require e-TeX (for example etoolbox): the e-TeX
%% extensions are not currently available on the ACS conversion
%% servers.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage[version=3]{mhchem} % Formula subscripts using \ce{}
\usepackage[obeyFinal]{easy-todo}
\usepackage{rotating}
\usepackage{epstopdf}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% If issues arise when submitting your manuscript, you may want to
%% un-comment the next line.  This provides information on the
%% version of every file you have used.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%\listfiles

@@

\section{The effect of cholesterol on glycerol backbone structure}
\begin{figure*}[]
  \centering
  \includegraphics[width=17.2cm]{../Fig/dihsCHOLcharmm.pdf}
  \caption{\label{dihsCHOLcharmm}
    The effect of cholesterol content on the POPC glycerol backbone and choline dihedral angles in CHARMM36 model (T=303~K).}
\end{figure*}

\begin{figure*}[]
  \centering
  \includegraphics[height=17.2cm]{../Fig/GlycerolOrientation.eps}
  \caption{\label{GlycerolOrientation}
    Orientation of g$_1$-g$_3$ (a) and g$_1$-g$_2$ (b) respect to membrane normal as a function of cholesterol content
    calculated from CHARMM36 simulations in 298K and 303K.}
\end{figure*}

\newpage

### 2015-08-12 HGmodelMANUSCRIPT/HGmodel_ACStemplate.tex
Updating the answer to reviewer's point 1 according to issue https://github.com/NMRLipids/nmrlipids.blogspot.fi/issues/57
Add glycerol orientation analysis and discussion to the manuscript, related to the reviewer's point 6

\subsection{Simulated systems}
All simulations are ran with a standard setup for a planar lipid bilayer in zero tension and constant temperature
with periodic boundary conditions in all directions by using the GROMACS software package~\cite{hess08} 
(version numbers 4.5.X--4.6.X), LAMMPS~\cite{plimpton95}, MDynaMix~\cite{Lyubartsev00} or NAMD~\cite{phillips05}.
The number of molecules, simulation temperatures and the length of simulations of all the simulated systems 
are listed in Tables~\ref{systems},~\ref{systemsDEHYD} and~\ref{systemsCHOL}. Full simulation
details are given in the Supporting Information (SI) or in the original publications in case the
data is used previously therein. Additionally, the files related to the simulations and the resulting trajectories are publicly
available for almost all systems in the Zenodo collection \url{https://zenodo.org/collection/user-nmrlipids}. 
The references pointing to simulation details and files are also listed in Tables~\ref{systems},~\ref{systemsDEHYD} and~\ref{systemsCHOL}.

@@

In addition to the experimental data, Fig.~\ref{ordPchol} shows
our results for the CHARMM36 and MacRog force fields
and the previously published~\cite{ferreira13}
Berger-POPC-07/H\"oltje-CHOL-13 results.
Note that the simulation data are shifted vertically to ease comparison with experimental responses.
As previously pointed out~\cite{ferreira13}, the Berger-based model
seriously exaggerates the effect of cholesterol on the phospholipid glycerol backbone and choline headgroup.
In comparison, the choline and glycerol backbone responses of CHARMM36 and MacRog are in better qualitative 
agreement with experiments. 
Therefore, to resolve the nature of cholesterol-induced structural changes,
we calculated from CHARMM36 the glycerol backbone dihedral angle distributions
at various cholesterol contents (Supporting material): The only detectable change
is the small decrease of gauche(-) and increase of gaughe(+) probability of the g$_3$-g$_2$-g$_1$-O(\textit{sn}-1) dihedral.
In conclusion, our results suggest that the significant effects of cholesterol on lipid conformations observed in simulations~\cite{zhu07,rog09,alwarawrah12} are overestimated by the computational models used;
cholesterol only induces very small structural changes in the glycerol backbone.

>>>>

\subsection{Simulated systems}
All simulations are ran with a standard setup for a planar lipid bilayer in zero tension and constant temperature
with periodic boundary conditions in all directions by using the GROMACS software package~\cite{hess08} 
(version numbers 4.5.X--4.6.X), LAMMPS~\cite{plimpton95}, MDynaMix~\cite{Lyubartsev00} or NAMD~\cite{phillips05}.
The number of molecules, simulation temperatures and the length of simulations of all the simulated systems 
are listed in Tables~\ref{systems},~\ref{systemsDEHYD} and~\ref{systemsCHOL}. Full simulation
details are given in the Supporting Information (SI) or in the original publications in case the
data is used previously therein. All simulation parameters were set as close to the original 
parametrization works as possible.
Additionally, the files related to the simulations and the resulting trajectories are publicly
available for almost all systems in the Zenodo collection \url{https://zenodo.org/collection/user-nmrlipids}. 
The references pointing to simulation details and files are also listed in Tables~\ref{systems},~\ref{systemsDEHYD} and~\ref{systemsCHOL}.

@@

In addition to the experimental data, Fig.~\ref{ordPchol} shows
our results for the CHARMM36 and MacRog force fields
and the previously published~\cite{ferreira13}
Berger-POPC-07/H\"oltje-CHOL-13 results.
Note that the simulation data are shifted vertically to ease comparison with experimental responses.
As previously pointed out~\cite{ferreira13}, the Berger-based model
seriously exaggerates the effect of cholesterol on the phospholipid glycerol backbone and choline headgroup.
In comparison, the choline and glycerol backbone responses of CHARMM36 and MacRog are in better qualitative 
agreement with experiments. 
Therefore, to resolve the nature of cholesterol-induced structural changes,
we calculated from CHARMM36 the glycerol backbone orientation and dihedral angle distributions
at various cholesterol contents (Supporting material). The only detectable changes
are the small decrease of gauche(-) and increase of gaughe(+) probability of 
the g$_3$-g$_2$-g$_1$-O(\textit{sn}-1) dihedral and slight (less than 5 degrees) change
in the glycerol backbone orientation. In conclusion, our results suggest that the significant 
effects of cholesterol on lipid conformations observed in simulations~\cite{zhu07,rog09,alwarawrah12} 
are overestimated by the computational models used;
cholesterol only induces very small structural changes in the glycerol backbone.
